Lattices of acyclic pipe dreams

### 2023-03-09 acyclicFacets/draft/latticePipeDreams.tex
adding quesiton
My reading comments and little fix.

\begin{figure}[ht]
	\centerline{\includegraphics[scale=1.2]{counterExampleRestrictionIncreasingFlipPoset}}
	\caption{Two acyclic pipe dreams~$\acyclicPipeDreams(126543)$ connected by a sequence of increasing flips in~$\pipeDreams(126543)$ but by no sequence of increasing flips in~$\acyclicPipeDreams(126543)$.}
	\label{fig:counterExampleRestrictionIncreasingFlipPoset}
\end{figure}
\end{remark}

@@

The \defn{weak order} on $W$ is the partial order $\le$ defined by $\sigma \leq \omega$ if there exists $\tau \in W$ such that $\sigma\tau = \omega$ and $\ell(\sigma) + \ell(\tau) = \ell(\omega)$.
In other words, the element $\omega$ has a reduced expression with a prefix which is a reduced expression of $\sigma$.
Equivalently, the weak order corresponds to the inclusion order on inversion sets, that is $\sigma \leq \omega$ if and only if $\Inv(\sigma) \subseteq \Inv(\omega)$.
This order defines a lattice structure on the elements of~$W$.
The minimal element is the identity $e \in W$ and the maximal element is the \defn{unique longest element} $\wo$ of~$W$.
Note that~$\Inv(e) = \varnothing = \Ninv(\wo)$ and~$\Ninv(e) = \Phi^+ = \Inv(\wo)$, so that~$\ell(e) = 0$ while~$\ell(\wo) = |\Phi^+|$.
% Note also that~$\Ninv(\omega) =\Inv(\omega\wo)$.

@@

\begin{remark}
\label{rem:linearExtensionsPartitionSubwordComplexA}
Before proving this theorem, we would like to make the following remarks about items~\eqref{item:convex} and~\eqref{item:cover}:
\begin{enumerate}
\setcounter{enumi}{0}
\item There are cases where $[e,\omega] \ne \linearExtensions(Q,\omega)$. Namely, some of the sets~$\linearExtensions(I)$ may not be included in~$[e,\omega]$, and may even not meet~$[e,\omega]$. See \cref{fig:counterExample1}.
%
%\begin{figure}[h]
%	\includegraphics[width=0.52\textwidth]{interval_not_contained1} \quad
%	\includegraphics[width=0.43\textwidth]{interval_not_contained2}
%	\caption{A facet of a subword complex~$\subwordComplex(Q, \omega)$ whose linear extensions all lie outside the interval~$[e, \omega]$.}
%	\label{fig:outOfInterval}
%\end{figure}
%
\begin{figure}[p]
	\centerline{\includegraphics[scale=.8]{counterExample1}}
	\caption{The sets~$\linearExtensions(I)$ (blue bubbles) are not always subsets of, and may even not meet~$[e,\omega]$ (green bubble).}
	\label{fig:counterExample1}
\end{figure}

\setcounter{enumi}{2}
\item Some of the sets~$\linearExtensions(I)$ may not be intervals. See \cref{fig:counterExample2}. Another example with~$\omega = \omega_\circ$ can be found in~\cite[Figure~9]{PilaudStump-brickPolytope}. 
%
\begin{figure}[p]
	\centerline{\includegraphics[scale=.8]{counterExample2}}
	\caption{The sets~$\linearExtensions(I)$ (blue bulles) are not always intervals.}
	\label{fig:counterExample2}
\end{figure}
\end{enumerate}
\cref{fig:counterExample1,fig:counterExample2} represent type~$A$ subword complexes.
As in~\cite{PilaudPocchiola, PilaudSantos-brickPolytope, PilaudStump-brickPolytope}, we thus use pseudoline arrangements on sorting networks to represent their facets (and we have represented their contact graphs together).
\end{remark}

@@

@@

@@

For the purpose of this proof, let us write~$\leqslant_B$ the strong Bruhat order, defined by~$x \leqslant_B y$ if and only if a reduced expression of~$y$ has a reduced expression of~$x$ as a subword, and~$x \prec_B y$ the covers of this order. We note that for any word~$Q$ and any~$x\in W$, we can find a reduced expression of~$x$ as a subword of~$Q$ if and only if~$x \leqslant_B Dem(Q)$ the demazure product of~$Q$.

We will use \cite[Proposition 3.14]{JahnStump}, reformulated as follows:

\begin{proposition}[\cite{JahnStump}]\label{prop:JSDemazureCone}
Let~$\subwordComplex(Q,w)$ be a non-empty subword complex. Then for any facet~$I \in \subwordComplex(Q,w)$ and for any simple root~$\alpha$ such that~$w \prec_B s_\alpha w \leqslant_B Dem(Q)$, we have~$\alpha \in \Roots(I)$.
\end{proposition}

Let us write~$w = uv$ with~$u$ the prefix of~$w$ defined by~$I_1$ and~$I_2$ on~$Q$ until index~$j-1$ and~$v$ the suffix of~$w$ defined by the same facets from index~$j$ and onward. Let us also write~$s$ the~$j^{\text{th}}$ letter of~$Q$ and~$\alpha$ the simple root associated with~$s$. We note that~$\beta = \rootFunction{I_1}{j} = u(\alpha)$, and that since~$j \notin I_2$, we know that~$sv \prec_B v$ (this is also a cover of the left weak order).

Let us call~$Q'$ the suffix of~$Q$ starting at index~$j+1$, and~$I'_1,I'_2$ the restrictions of~$I_1$ and~$I_2$ to~$Q'$. The subwords induced by~$I'_1$ and~$I'_2$ on~$Q'$ are suffixes of reduced subwords of~$Q$, so they are also reduced. Moreover, since~$j \in I_1$, we know that~$I'_1$ is a facet of~$\subwordComplex(Q',v)$; this means that this subword complex is not empty and that~$v \leqslant_B Dem(Q')$. Similarly, since~$j \notin I_2$, we know that~$I'_2$ is a facet of~$\subwordComplex(Q',sv)$. By combining the previous statements, we know that~$sv \prec_B v \leqslant_B Dem(Q')$ (or equivalently and with~$v' = sv$, that~$v' \prec_B sv' \leqslant_B Dem(Q')$). We can thus apply \cref{prop:JSDemazureCone} to obtain that~$\alpha \in \Roots(I'_2)$.

Going back to our facets on~$Q$, we know that the prefix of~$w$ written by~$I_2$ on~$Q \cap [j]$ is~$us$, and thus~$us(\Roots(I'_2)) \subseteq \Roots(I_2)$. Therefore, we have~$us(\alpha) = u(-\alpha) = -u(\alpha) = -\beta \in \Roots(I_2)$, thus concluding the proof.
\end{proof}

>>>>

\begin{figure}[ht]
	\centerline{\includegraphics[scale=1.3]{counterExampleRestrictionIncreasingFlipPoset}}
	\caption{Two acyclic pipe dreams~$\acyclicPipeDreams(126543)$ connected by a sequence of increasing flips in~$\pipeDreams(126543)$ but by no sequence of increasing flips in~$\acyclicPipeDreams(126543)$.}
	\label{fig:counterExampleRestrictionIncreasingFlipPoset}
\end{figure}
\end{remark}

@@

The \defn{weak order} on $W$ is the partial order $\le$ defined by $\sigma \leq \omega$ if there exists $\tau \in S$  such that $\sigma\tau = \omega$ and $\ell(\sigma) + \ell(\tau) = \ell(\omega)$.
In other words, the element $\omega$ has a reduced expression with a prefix which is a reduced expression of $\sigma$.
Equivalently, the weak order corresponds to the inclusion order on inversion sets, that is $\sigma \leq \omega$ if and only if $\Inv(\sigma) \subseteq \Inv(\omega)$.
This order defines a lattice structure on the elements of~$W$.
The minimal element is the identity $e \in W$ and the maximal element is the \defn{unique longest element} $\wo$ of~$W$.
Note that~$\Inv(e) = \varnothing = \Ninv(\wo)$ and~$\Ninv(e) = \Phi^+ = \Inv(\wo)$, so that~$\ell(e) = 0$ while~$\ell(\wo) = |\Phi^+|$.
% Note also that~$\Ninv(\omega) =\Inv(\omega\wo)$.

@@

\begin{remark}
\label{rem:linearExtensionsPartitionSubwordComplexA}
Before proving this theorem, we would like to make the following remarks about items~\eqref{item:convex} and~\eqref{item:cover}.
\begin{enumerate}
\setcounter{enumi}{0}
\item There are cases where $[e,\omega] \ne \linearExtensions(Q,\omega)$. \cesarm{add discussion}
%
\begin{figure}[h]
	\includegraphics[width=0.52\textwidth]{interval_not_contained1} \quad
	\includegraphics[width=0.43\textwidth]{interval_not_contained2}
	\caption{A facet of a subword complex~$\subwordComplex(Q, \omega)$ whose linear extensions all lie outside the interval~$[e, \omega]$.}
	\label{fig:outOfInterval}
\end{figure}
%
\setcounter{enumi}{2}
\item $\linearExtensions(I)$ is not an interval in general. See for instance the example~\cite[Figure~9]{PilaudStump-brickPolytope}. 
In that figure, the elements of the week order $\mathfrak{S}_4$ that are linear extensions of the same acyclic facet of a given subword complex are grouped together. Some of the groups in the figure are not intervals.  \\
\end{enumerate}
\end{remark}

@@

\nantel{it is bad to name this ``congruence". Call it equivalence, and state it is a concruence in the conjecture.}

@@

\nantel{to be honest, I find that the paper ends well here, the rest should be part of a paper by No\'emie alone?}

@@

For the purpose of this proof, let us write~$\leqslant_B$ the strong Bruhat order, defined by~$x \leqslant_B y \iff$ a reduced expression of~$y$ has a reduced expression of~$x$ as a subword, and~$x \less_B y$ the covers of this order. We note that for any word~$Q$ and any~$x\in W$, we can find a reduced expression of~$x$ as a subword of~$Q$ if and only if~$x \leqslant_B Dem(Q)$ the demazure product of~$Q$.

Let us write~$w = uv$ with~$u$ the prefix of~$w$ defined by~$I_1$ and~$I_2$ on~$Q$ until index~$j-1$ and~$v$ the suffix of~$w$ defined by the same facets from index~$j$ and onward. Let us also write~$s$ the~$j^{\text{th}}$ letter of~$Q$ and~$\alpha$ the simple root associated with~$s$. We note that~$\beta = \rootFunction{I_1}{j} = u(\alpha)$.

Let us call~$Q'$ the suffix of~$Q$ starting at index~$j+1$. Since~$j \notin I_2$, we note that~$sv \less_B v$ and the restriction of~$I_2$ to~$Q'$ is a facet of~$\subwordComplex(Q', sv)$. However, since~$j \in I_1$, the restriction of~$I_1$ to~$Q'$ is a facet of~$\subwordComplex(Q', v)$. Therefore, we know that~$v \leqslant_B Dem(Q')$.

We can now apply \cite[proposition 3.14]{JahnStump} to~$I'_2$ the restriction of~$I_2$ to~$Q'$, which tells us that~$\mathcal{E}^+(sv, Dem(Q)) \subseteq \cone \Roots(I'_2)$. Since~$sv \less_B v \leqslant Dem(Q)$, then~$\alpha \in \mathcal{E}^+(sv, Dem(Q))$ and so~$\alpha \in \cone \Roots(I'_2)$.

Going back to~$I_2$, we know that~$us(\Roots(I'_2)) \subseteq \Roots(I_2)$ and so~$us(\alpha) = -u(\alpha) = -\beta \in \cone \Roots(I_2)$.
\end{proof}
